---
title: "Exercise 3.1.6"
description: | 
  Individual exercise document
author: |
  | [Saurabh Steixner-Kumar](https://steixnerkumar.github.io/)
  | \footnotesize (social): [\large\faLinkedinSquare](https://www.linkedin.com/in/SteixnerKumar/) -  [\large\faGithub](https://github.com/SteixnerKumar) -  [\large\faTwitter](https://twitter.com/SteixnerKumar)
output: 
  html_document: 
    fig_caption: yes
    fig_height: 10
    fig_width: 17
    number_sections: yes
    toc: yes
    toc_float: true
  html_notebook: 
    fig_caption: yes
    fig_height: 10
    fig_width: 17
    number_sections: yes
    toc: yes
  pdf_document:
    latex_engine: xelatex 
    fig_caption: yes
    fig_height: 6
    fig_width: 15
    keep_tex: no
    number_sections: yes
    toc: yes
editor_options: 
  chunk_output_type: inline
# runtime: shiny
header-includes:
  - \usepackage{titling}
  - '\defaultfontfeatures{Extension = .otf}'
  - '\usepackage{fontawesome}'
  - \preauthor{\begin{flushright}}
  - \postauthor{\end{flushright}}
---

# Question
Exercise 3.1.6 Alter the data to k = 0 and n = 1, and comment on the shape
of the posterior for the rate $\theta$.

# Comments/Solution
Comparing at the posterior distributions (from the plots section below) of k=0, n=1 to the previous distribution of k=50, n=100, one can observe that the $\theta$ estimates have shifted very slightly to the lekt and towards zero from the prior value of a uniform distribution. The shift is very subtle because the data sample is only 1. This shift would become more prominent like exercise 3.1.5 with higher n towards left.

The model used to calculate the required values and the plots is scripted below. Copy/pasting the given code will generate the same result on your own machine.


# Code
## libraries
The libraries required for the script and the plots.
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
# clears workspace
rm(list=ls()) 
#load libraries
library(rstan)
library(ggplot2)
library(patchwork)
```


## Data
The data required for this particular stan model.
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
# data  initialization
k <- 50
n <- 100
k_1 <- 0
n_1 <- 1
# to be passed on to Stan
stan_data <- list(k = k, n = n)
stan_data_1 <- list(k = k_1, n = n_1)
```


## Stan code
Stan code, that can be written in R as such or in a seperate new file with stan extension.
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
write("// Stan code here in this section

// Inferring theta
data { 
  int<lower=1> n; 
  int<lower=0> k;
} 
parameters {
  real<lower=0,upper=1> theta;
} 
model {
  // Prior Distribution for theta
  theta ~ beta(1, 1);
  
  // Observed Counts
  k ~ binomial(n, theta);
} // ",

"3_1_6.stan")
```

## code in R to run stan
Running stan through R (with the required input parameters).
```{r echo=TRUE, results = 'hide', message=FALSE, warning=FALSE, paged.print=TRUE}

myinits <- list(
  list(theta=.1),  # chain 1 starting value
  list(theta=.9))  # chain 2 starting value

# parameters to be monitored:  
parameters <- c("theta")

# The following command calls Stan with specific options.
# For a detailed description type "?stan".
mod_fit <- stan(file="3_1_6.stan",   
                data=stan_data, 
                init=myinits,  # If not specified, gives random inits
                pars=parameters,
                iter=2000, 
                chains=2, 
                thin=1,
                warmup=100,  # Stands for burn-in; Default = iter/2
                seed=123  # Setting seed; Default is random seed
)
mod_fit_1 <- stan(file="3_1_6.stan",   
                data=stan_data_1, 
                init=myinits,  # If not specified, gives random inits
                pars=parameters,
                iter=2000, 
                chains=2, 
                thin=1,
                warmup=100,  # Stands for burn-in; Default = iter/2
                seed=123  # Setting seed; Default is random seed
)
```

# Outputs
## Model summary
For dataset k=50 and n=100. (stan_data)
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# The commands below are useful for a quick overview:
print(mod_fit)  # a rough summary
#plot(mod_fit)   # a visual representation
```
For dataset k=0 and n=1. (stan_data_1)
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# The commands below are useful for a quick overview:
print(mod_fit_1)  # a rough summary
```

## Plots

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
# Diagnostic plot
# stan_par(mod_fit, 'theta', chain = 0)

# First, some options to make the plot look better:
#
plot_color_1 <- adjustcolor("black",alpha.f=1)
plot_color_2 <- adjustcolor("white",alpha.f=1)
plot_color_3 <- adjustcolor("darkred",alpha.f=1)
plot_color_4 <- adjustcolor("darkgreen",alpha.f=1)
plot_color_5 <- adjustcolor("blue",alpha.f=0.8)
plot_text_size <- 32
plot_text_color <- adjustcolor("black",alpha.f=1)

```


### Plot (chains)
The initial movement of the chains are shown here (including the warmup phase). The two chains begin from the initial starting points of as defined in the input parameters of the stan model.
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# chain plot
x_limit <- c(0,200)
trace_1 <- traceplot(mod_fit, 'theta', include = TRUE, unconstrain = FALSE,inc_warmup = TRUE)+
  xlim(x_limit[1], x_limit[2])+
  scale_color_manual(values = c(plot_color_3, plot_color_4)) +
  theme(legend.position = "bottom")+
  theme_classic()+
  ylab('theta')+xlab('')+ggtitle(paste0('k=',stan_data$k,', n=',stan_data$n))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y = element_text(color = plot_text_color, size = plot_text_size))+
  theme(plot.title = element_text(hjust=0.5))
trace_2 <- traceplot(mod_fit_1, 'theta', include = TRUE, unconstrain = FALSE,inc_warmup = TRUE)+
  xlim(x_limit[1], x_limit[2])+
  scale_color_manual(values = c(plot_color_3, plot_color_4)) +
  theme(legend.position = "bottom")+
  theme_classic()+
  ylab('theta')+xlab('iterations')+ggtitle(paste0('k=',stan_data_1$k,', n=',stan_data_1$n))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y = element_text(color = plot_text_color, size = plot_text_size))+
  theme(plot.title = element_text(hjust=0.5))
patchwork <- trace_1/trace_2
p_ranges_x <- c(ggplot_build(patchwork[[1]])$layout$panel_scales_x[[1]]$range$range,
  ggplot_build(patchwork[[2]])$layout$panel_scales_x[[1]]$range$range)

p_ranges_y <- c(ggplot_build(patchwork[[1]])$layout$panel_scales_y[[1]]$range$range,
                ggplot_build(patchwork[[2]])$layout$panel_scales_y[[1]]$range$range)

patchwork + plot_annotation(title = 'comparison of datasets') +
  plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = plot_text_size+2),
                                            plot.title=element_text(size = plot_text_size+2,hjust = 0.5)) &
  xlim(min(p_ranges_x), max(p_ranges_x)) & 
  ylim(min(p_ranges_y), max(p_ranges_y))
```

### Plot (posterior)
The plot of the $\theta$ values per chain superimposed on each other.
```{r, fig.align="center", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# Now let's plot a histogram/density for theta.
# Plot

dens_1 <- stan_dens(mod_fit, 'theta', alpha = 0.3, include = TRUE, unconstrain = FALSE,
            inc_warmup = FALSE, separate_chains = TRUE,show.legend = FALSE) +
  scale_fill_manual(values = c(plot_color_3, plot_color_4))+
  xlab('theta')+ylab('density')+ggtitle(paste0('k=',stan_data$k,', n=',stan_data$n))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y = element_text(color = plot_text_color, angle = 90,size = plot_text_size))+
  theme(plot.title = element_text(hjust=0.5))
dens_2 <- stan_dens(mod_fit_1, 'theta', alpha = 0.3, include = TRUE, unconstrain = FALSE,
            inc_warmup = FALSE, separate_chains = TRUE) +
  scale_fill_manual(values = c(plot_color_3, plot_color_4))+
  xlab('theta')+ylab('density')+ggtitle(paste0('k=',stan_data_1$k,', n=',stan_data_1$n))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y =element_blank())+
  theme(plot.title = element_text(hjust=0.5))
patchwork <- dens_1+dens_2
p_ranges_x <- c(ggplot_build(patchwork[[1]])$layout$panel_scales_x[[1]]$range$range,
  ggplot_build(patchwork[[2]])$layout$panel_scales_x[[1]]$range$range)

p_ranges_y <- c(ggplot_build(patchwork[[1]])$layout$panel_scales_y[[1]]$range$range,
                ggplot_build(patchwork[[2]])$layout$panel_scales_y[[1]]$range$range)

patchwork + plot_annotation(title = 'comparison of datasets') +
  plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = plot_text_size+2),
                                            plot.title=element_text(size = plot_text_size+2,hjust = 0.5)) &
  xlim(min(p_ranges_x), max(p_ranges_x)) & 
  ylim(min(p_ranges_y), max(p_ranges_y))

```



