---
title: "Exercise 3.1.2"
description: | 
  Individual exercise document
author: |
  | [Saurabh Steixner-Kumar](https://steixnerkumar.github.io/)
  | \footnotesize (social): [\large\faLinkedinSquare](https://www.linkedin.com/in/SteixnerKumar/) -  [\large\faGithub](https://github.com/SteixnerKumar) -  [\large\faTwitter](https://twitter.com/SteixnerKumar)
output: 
  html_document: 
    fig_caption: yes
    fig_height: 10
    fig_width: 17
    number_sections: yes
    toc: yes
    toc_float: true
  html_notebook: 
    fig_caption: yes
    fig_height: 10
    fig_width: 17
    number_sections: yes
    toc: yes
  pdf_document:
    latex_engine: xelatex 
    fig_caption: yes
    fig_height: 10
    fig_width: 17
    keep_tex: no
    number_sections: yes
    toc: yes
editor_options: 
  chunk_output_type: inline
# runtime: shiny
header-includes:
  - \usepackage{titling}
  - '\defaultfontfeatures{Extension = .otf}'
  - '\usepackage{fontawesome}'
  - \preauthor{\begin{flushright}}
  - \postauthor{\end{flushright}}
---

# Question
Exercise 3.1.2 Consider again the posterior distribution for $\theta$ given k = 5 successes
out of n = 10 trials. Based on a visual impression, what is your estimate
of how much more likely the rate $\theta$ is to equal to 0.5 rather than 0.7? How
did you arrive at your estimate?

# Comments/Solution
Looking at the posterior distribution (from the plots section below), one can estimate that the value of $\theta$ is approximately twice more likely to be 0.5 than 0.7. We can calculate the exact value from the estimates as:

```{r echo=TRUE, eval = FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
length(theta[theta>0.49 & theta<0.51])/length(theta[theta>0.69 & theta<0.71])
```
The value is given in the last section of this page under 'calculation'.
The model used to calculate the required values and the plots is scripted below. Copy/pasting the given code will generate the same result on your own machine.


# Code
## libraries
The libraries required for the script and the plots.
```{r libraries, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
# clears workspace
rm(list=ls()) 
#load libraries
library(rstan)
library(ggplot2)
library(patchwork)
```


## Data
The data required for this particular stan model.
```{r data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
# data  initialization
k <- 5
n <- 10
# to be passed on to Stan
stan_data <- list(k = k, n = n)
```


## Stan code
Stan code, that can be written in R as such or in a separate new file with stan extension.
```{r stan_code, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
write("// Stan code here in this section

// Inferring theta
data { 
  int<lower=1> n; 
  int<lower=0> k;
} 
parameters {
  real<lower=0,upper=1> theta;
} 
model {
  // Prior Distribution for theta
  theta ~ beta(1, 1);
  
  // Observed Counts
  k ~ binomial(n, theta);
} // ",

"3_1_2.stan")
```

## code in R to run stan
Running stan through R (with the required input parameters).
```{r running_stan, echo=TRUE, results = 'hide', message=FALSE, warning=FALSE, paged.print=TRUE}

myinits <- list(
  list(theta=.1),  # chain 1 starting value
  list(theta=.9))  # chain 2 starting value

# parameters to be monitored:  
parameters <- c("theta")

# The following command calls Stan with specific options.
# For a detailed description type "?stan".
mod_fit <- stan(file="3_1_2.stan",   
                data=stan_data, 
                init=myinits,  # If not specified, gives random inits
                pars=parameters,
                iter=2000, 
                chains=2, 
                thin=1,
                warmup=100,  # Stands for burn-in; Default = iter/2
                seed=123  # Setting seed; Default is random seed
)
```

# Outputs
## Model summary
In order of definition.
```{r model_summary, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# The commands below are useful for a quick overview:
print(mod_fit)  # a rough summary
#plot(mod_fit)   # a visual representation

# extract the data from the model
mod_fit_extracted <- rstan::extract(mod_fit,inc_warmup = FALSE)
loc_param <- which(names(mod_fit_extracted) == 'theta')
param <- as.array(mod_fit)[,,loc_param]

# Collect posterior and prior samples across all chains:
theta <- mod_fit_extracted$theta
```

## Plots

```{r plot_defaults, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
# Diagnostic plot
# stan_par(mod_fit, 'theta', chain = 0)

# First, some options to make the plot look better:
#
plot_color_1 <- adjustcolor("black",alpha.f=1)
plot_color_2 <- adjustcolor("white",alpha.f=1)
plot_color_3 <- adjustcolor("darkred",alpha.f=1)
plot_color_4 <- adjustcolor("darkgreen",alpha.f=1)
plot_color_5 <- adjustcolor("blue",alpha.f=0.8)
plot_text_size <- 32
plot_text_color <- adjustcolor("black",alpha.f=1)

```


### Plot (chains)
The initial movement of the chains are shown here (including the warmup phase). The two chains begin from the initial starting points of as defined in the input parameters of the stan model.
```{r plot_chain, fig.align="center", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# chain plot
traceplot(mod_fit, 'theta', include = TRUE, unconstrain = FALSE,inc_warmup = TRUE)+
  xlim(0, 200)+
  scale_color_manual(values = c(plot_color_3, plot_color_4)) +
  theme(legend.position = "bottom")+
  theme_classic()+
  ylab('theta')+xlab('iterations')+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y = element_text(color = plot_text_color, size = plot_text_size))+
  theme(plot.title = element_text(hjust=0.5))

```

### Plot (posterior)
The plot of the $\theta$ values per chain superimposed on each other.
```{r plot_posterior, fig.align="center", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# Now let's plot a histogram/density for theta.
# Plot

stan_dens(mod_fit, 'theta', alpha = 0.3, include = TRUE, unconstrain = FALSE,
            inc_warmup = FALSE, separate_chains = TRUE) +
  scale_fill_manual(values = c(plot_color_3, plot_color_4))+
  xlab('theta')+ylab('density')+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y = element_text(color = plot_text_color, angle = 90,size = plot_text_size))+
  theme(plot.title = element_text(hjust=0.5))

```

# Calculation
Calculation from the comment/solution.
```{r extra_calculation, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
length(theta[theta>0.49 & theta<0.51])/length(theta[theta>0.69 & theta<0.71])
```


