---
title: "Exercise 3.3.1"
description: | 
  Individual exercise document
author: |
  | [Saurabh Steixner-Kumar](https://steixnerkumar.github.io/)
  | \footnotesize (social): [\large\faLinkedinSquare](https://www.linkedin.com/in/SteixnerKumar/) -  [\large\faGithub](https://github.com/SteixnerKumar) -  [\large\faTwitter](https://twitter.com/SteixnerKumar)
output: 
  html_document: 
    fig_caption: yes
    fig_height: 10
    fig_width: 17
    number_sections: yes
    toc: yes
    toc_float: true
  html_notebook: 
    fig_caption: yes
    fig_height: 10
    fig_width: 17
    number_sections: yes
    toc: yes
  pdf_document:
    latex_engine: xelatex 
    fig_caption: yes
    fig_height: 10
    fig_width: 17
    keep_tex: no
    number_sections: yes
    toc: yes
editor_options: 
  chunk_output_type: inline
# runtime: shiny
header-includes:
  - \usepackage{titling}
  - '\defaultfontfeatures{Extension = .otf}'
  - '\usepackage{fontawesome}'
  - \preauthor{\begin{flushright}}
  - \postauthor{\end{flushright}}
---

# Question
Exercise 3.3.1 Try the data k1 = 14, n1 = 20, k2 = 16, n2 = 20. How could you
report the inference about the common rate $\theta$?

# Comments/Solution
In this example of the dataset we see that the individually the k1 and k2 both present cases around the probability of 0.75 and the common rate $\theta$ is estimated also to be around 0.75. Check out the plots section.

One reasonable reporting strategy here might be to use a measure for central tendency, such as a mean, median, or mode, together with a credible interval, for instance a 95% credible interval.

The model used to calculate the required values and the plots is scripted below. Copy/pasting the given code will generate the same result on your own machine.


# Code
## libraries
The libraries required for the script and the plots.
```{r libraries, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
# clears workspace
rm(list=ls()) 
#load libraries
library(rstan)
library(bayestestR)
library(ggplot2)
library(patchwork)
```


## Data
The data required for this particular stan model.
```{r data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
# data  initialization
k1 <- 14;n1 <- 20;k2 <- 16;n2 <- 20
# to be passed on to Stan
stan_data <- list(k1 = k1, n1 = n1, k2 = k2, n2 = n2)
```


## Stan code
Stan code, that can be written in R as such or in a separate new file with stan extension.
```{r stan_code, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
write("// Stan code here in this section

// Inferring theta
data { 
  int<lower=1> n1; 
  int<lower=1> n2; 
  int<lower=0> k1;
  int<lower=0> k2;
} 
parameters {
  real<lower=0,upper=1> theta;
} 
model {
  // Prior on Single Rate Theta
  theta ~ beta(1, 1);
  // Observed Counts
  k1 ~ binomial(n1, theta);
  k2 ~ binomial(n2, theta);
} // ",

"3_3_1.stan")
```

## code in R to run stan
Running stan through R (with the required input parameters).
```{r running_stan, echo=TRUE, results = 'hide', message=FALSE, warning=FALSE, paged.print=TRUE}

myinits <- list(
  list(theta=.1),  # chain 1 starting value
  list(theta=.9))  # chain 2 starting value

# parameters to be monitored:  
parameters <- c("theta")

# The following command calls Stan with specific options.
# For a detailed description type "?stan".
mod_fit <- stan(file="3_3_1.stan",   
                data=stan_data, 
                init=myinits,  # If not specified, gives random inits
                pars=parameters,
                iter=2000, 
                chains=2, 
                thin=1,
                warmup=100,  # Stands for burn-in; Default = iter/2
                seed=123  # Setting seed; Default is random seed
)
```

# Outputs
## Model summary
In order of definition.
```{r model_summary, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# The commands below are useful for a quick overview:
print(mod_fit)  # a rough summary
#plot(mod_fit)   # a visual representation
```

```{r model_parameters, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
# get the data from the model
theta   <- rstan::extract(mod_fit)$theta
#
# get the confidence interval
ci_hdi <- ci(theta, ci=0.95,method = "HDI")
ci_eti <- ci(theta, ci=0.95,method = "ETI")
```

## Plots

```{r plot_defaults, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
# Diagnostic plot
# stan_par(mod_fit, 'theta', chain = 0)

# First, some options to make the plot look better:
#
plot_color_1 <- adjustcolor("black",alpha.f=1)
plot_color_2 <- adjustcolor("white",alpha.f=1)
plot_color_3 <- adjustcolor("darkred",alpha.f=1)
plot_color_4 <- adjustcolor("darkgreen",alpha.f=1)
plot_color_5 <- adjustcolor("blue",alpha.f=0.8)
plot_text_size <- 32
plot_text_color <- adjustcolor("black",alpha.f=1)

```


### Plot (chains)
The initial movement of the chains are shown here (including the warmup phase). The two chains begin from the initial starting points of as defined in the input parameters of the stan model.
```{r plot_chain, fig.align="center", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# chain plot
x_limit <- c(0,200)
trace_1 <- traceplot(mod_fit, 'theta', include = TRUE, unconstrain = FALSE,inc_warmup = TRUE,show.legend = FALSE)+
  xlim(x_limit[1], x_limit[2])+
  scale_color_manual(values = c(plot_color_3, plot_color_4)) +
  theme(legend.position = "bottom")+
  theme_classic()+
  ylab('theta')+xlab('iterations')+
  ggtitle(paste0('k1=',stan_data$k1,', n1=',stan_data$n1,'\n','k2=',stan_data$k2,', n2=',stan_data$n2))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y = element_text(color = plot_text_color, size = plot_text_size))+
  theme(plot.title = element_text(hjust=0.5))
trace_1
```

### Plot (posterior)
The plot of the $\theta$ values per chain superimposed on each other.
```{r plot_posterior, fig.align="center", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# Now let's plot a histogram/density for theta.
# Plot

dens_1 <- stan_dens(mod_fit, 'theta', alpha = 0.3, include = TRUE, unconstrain = FALSE,
            inc_warmup = FALSE, separate_chains = TRUE,show.legend = FALSE) +
  scale_fill_manual(values = c(plot_color_3, plot_color_4))+
  xlab('theta')+ylab('density')+
  ggtitle(paste0('k1=',stan_data$k1,', n1=',stan_data$n1,'\n','k2=',stan_data$k2,', n2=',stan_data$n2))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y = element_text(color = plot_text_color, angle = 90,size = plot_text_size))+
  theme(plot.title = element_text(hjust=0.5))
dens_1
```



## Confidence interval value and the measures of central tendency
The confidence interval values.
```{r ci, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE, paged.print=TRUE}
# Confidence intervals
ci_hdi
ci_eti
```
The mean, median and mode values respectively
```{r mean_median_mode, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE, paged.print=TRUE}
# function to calculate mode
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
#
c(mean(theta),median(theta),getmode(theta))
```