---
title: "Exercise 3.4.1"
description: | 
  Individual exercise document
author: |
  | [Saurabh Steixner-Kumar](https://steixnerkumar.github.io/)
  | \footnotesize (social): [\large\faLinkedinSquare](https://www.linkedin.com/in/SteixnerKumar/) -  [\large\faGithub](https://github.com/SteixnerKumar) -  [\large\faTwitter](https://twitter.com/SteixnerKumar)
output: 
  html_document: 
    fig_caption: yes
    fig_height: 10
    fig_width: 17
    number_sections: yes
    toc: yes
    toc_float: true
  html_notebook: 
    fig_caption: yes
    fig_height: 10
    fig_width: 17
    number_sections: yes
    toc: yes
  pdf_document:
    latex_engine: xelatex 
    fig_caption: yes
    fig_height: 10
    fig_width: 17
    keep_tex: no
    number_sections: yes
    toc: yes
editor_options: 
  chunk_output_type: inline
# runtime: shiny
header-includes:
  - \usepackage{titling}
  - '\defaultfontfeatures{Extension = .otf}'
  - '\usepackage{fontawesome}'
  - \preauthor{\begin{flushright}}
  - \postauthor{\end{flushright}}
---

# Question
Exercise 3.4.1 Make sure you understand the prior, posterior, prior predictive and posterior predictive distributions, and how they relate to each other (e.g., why is the top panel of Figure 3.9 a line plot, while the bottom panel is a bar graph?). Understanding these ideas is a key to understanding Bayesian analysis. Check your understanding by trying other data sets, varying both k and n.

# Comments/Solution
The top panel represents the density (positive hits) of the parameters while the bottom panel is the number of counts of the prior and posterior positive hits. This is based on the given data for 5 positive hit out of 15. Look at the plots section (below) to visually inspect it further. You will also see the trace chain for the theta_prior all over the place, showing a flat prior (density plot clears it further).

Line plots are for continuous quantities (e.g., rate parameter $\theta$) and bar plots are for discrete quantities (e.g., success counts of data).

The model used to calculate the required values and the plots is scripted below. Copy/pasting the given code will generate the same result on your own machine.


# Code
## libraries
The libraries required for the script and the plots.
```{r libraries, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
# clears workspace
rm(list=ls()) 
#load libraries
library(rstan)
library(tidyr)
library(ggplot2)
library(patchwork)
```


## Data
The data required for this particular stan model.
```{r data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
# data  initialization
k <- 5;n <- 15
# to be passed on to Stan
stan_data <- list(k = k, n = n)
```


## Stan code
Stan code, that can be written in R as such or in a separate new file with stan extension.
```{r stan_code, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
write("// Stan code here in this section

// Prior and Posterior Prediction
data {
  int<lower=1> n; 
  int<lower=0> k;
} 
parameters {
  real<lower=0,upper=1> theta;
  real<lower=0,upper=1> thetaprior;
}
model {
  // Prior on theta
  theta ~ beta(1, 1);
  thetaprior ~ beta(1, 1);
  // Observed Data
  k ~ binomial(n, theta);
}
generated quantities {
  int<lower=0> postpredk;
  int<lower=0> priorpredk;
    
  // Posterior Predictive
  postpredk = binomial_rng(n, theta);
  // Prior Predictive
  priorpredk = binomial_rng(n, thetaprior);
} // ",

"3_4_1.stan")
```

## code in R to run stan
Running stan through R (with the required input parameters).
```{r  running_stan, echo=TRUE, results = 'hide', message=FALSE, warning=FALSE, paged.print=TRUE}

myinits <- list(
  list(theta=.5,thetaprior=.5),  # chain 1 starting value
  list(theta=.5,thetaprior=.5))  # chain 2 starting value

# parameters to be monitored:  
parameters <- c("theta", "thetaprior", "postpredk", "priorpredk")

# The following command calls Stan with specific options.
# For a detailed description type "?stan".
mod_fit <- stan(file="3_4_1.stan",   
                data=stan_data, 
                init=myinits,  # If not specified, gives random inits
                pars=parameters,
                iter=2000, 
                chains=2, 
                thin=1,
                warmup=100,  # Stands for burn-in; Default = iter/2
                seed=123  # Setting seed; Default is random seed
)
```

# Outputs
## Model summary
In order of definition.
```{r model_summary, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# The commands below are useful for a quick overview:
print(mod_fit)  # a rough summary
#plot(mod_fit)   # a visual representation
```

```{r model_parameters, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
# get the data from the model
theta   <- rstan::extract(mod_fit)$theta
thetaprior  <- rstan::extract(mod_fit)$thetaprior 
priorpredk  <- rstan::extract(mod_fit)$priorpredk
postpredk   <- rstan::extract(mod_fit)$postpredk
```

## Plots

```{r plot_defaults, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
# Diagnostic plot
# stan_par(mod_fit, 'theta', chain = 0)

# First, some options to make the plot look better:
#
plot_color_1 <- adjustcolor("black",alpha.f=1)
plot_color_2 <- adjustcolor("white",alpha.f=1)
plot_color_3 <- adjustcolor("darkred",alpha.f=1)
plot_color_4 <- adjustcolor("darkgreen",alpha.f=1)
plot_color_5 <- adjustcolor("blue",alpha.f=0.8)
plot_color_6 <- adjustcolor("darkred",alpha.f=0.6)
plot_color_7 <- adjustcolor("darkgreen",alpha.f=0.6)
plot_text_size <- 32
plot_text_color <- adjustcolor("black",alpha.f=1)

```


### Plot (chains)
The initial movement of the chains are shown here (including the warmup phase). The two chains begin from the initial starting points of as defined in the input parameters of the stan model.
```{r plot_chain, fig.align="center", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# chain plot
x_limit <- c(0,200)
trace_1 <- traceplot(mod_fit, 'theta', include = TRUE, unconstrain = FALSE,inc_warmup = TRUE,show.legend = FALSE)+
  xlim(x_limit[1], x_limit[2])+
  scale_color_manual(values = c(plot_color_3, plot_color_4)) +
  theme(legend.position = "bottom")+
  theme_classic()+
  ylab('theta')+xlab('iterations')+
  ggtitle(paste0('k=',stan_data$k,', n=',stan_data$n))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y = element_text(color = plot_text_color, size = plot_text_size))+
  theme(plot.title = element_text(hjust=0.5))
trace_2 <- traceplot(mod_fit, 'thetaprior', include = TRUE, unconstrain = FALSE,inc_warmup = TRUE)+
  xlim(x_limit[1], x_limit[2])+
  scale_color_manual(values = c(plot_color_3, plot_color_4)) +
  theme(legend.position = "bottom")+
  theme_classic()+
  ylab('theta_prior')+xlab('iterations')+
  ggtitle(paste0('k=',stan_data$k,', n=',stan_data$n))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y = element_text(color = plot_text_color, size = plot_text_size))+
  theme(plot.title = element_text(hjust=0.5))
patchwork <- trace_1+trace_2
p_ranges_x <- c(ggplot_build(patchwork[[1]])$layout$panel_scales_x[[1]]$range$range,
  ggplot_build(patchwork[[2]])$layout$panel_scales_x[[1]]$range$range)

p_ranges_y <- c(ggplot_build(patchwork[[1]])$layout$panel_scales_y[[1]]$range$range,
                ggplot_build(patchwork[[2]])$layout$panel_scales_y[[1]]$range$range)

patchwork + plot_annotation(title = 'the two parameters') +
  plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = plot_text_size+2),
                                            plot.title=element_text(size = plot_text_size+2,hjust = 0.5)) &
  xlim(min(p_ranges_x), max(p_ranges_x)) & 
  ylim(min(p_ranges_y), max(p_ranges_y))

```

### Plot (posterior)
The plot of the $\theta$ values per chain superimposed on each other.
```{r plot_posterior_1, fig.align="center", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# Now let's plot a histogram/density for theta.
# Plot

dens_1 <- stan_dens(mod_fit, 'theta', alpha = 0.3, include = TRUE, unconstrain = FALSE,
            inc_warmup = FALSE, separate_chains = TRUE,show.legend = FALSE) +
  scale_fill_manual(values = c(plot_color_3, plot_color_4))+
  xlab('theta')+ylab('density')+
  ggtitle(paste0('k=',stan_data$k,', n=',stan_data$n))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y = element_text(color = plot_text_color, angle = 90,size = plot_text_size))+
  theme(plot.title = element_text(hjust=0.5))
dens_2 <- stan_dens(mod_fit, 'thetaprior', alpha = 0.3, include = TRUE, unconstrain = FALSE,
            inc_warmup = FALSE, separate_chains = TRUE) +
  scale_fill_manual(values = c(plot_color_3, plot_color_4))+
  xlab('theta_prior')+ylab('density')+
  ggtitle(paste0('k=',stan_data$k,', n=',stan_data$n))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y =element_blank())+
  theme(plot.title = element_text(hjust=0.5))
patchwork <- dens_1+dens_2
p_ranges_x <- c(ggplot_build(patchwork[[1]])$layout$panel_scales_x[[1]]$range$range,
  ggplot_build(patchwork[[2]])$layout$panel_scales_x[[1]]$range$range)

p_ranges_y <- c(ggplot_build(patchwork[[1]])$layout$panel_scales_y[[1]]$range$range,
                ggplot_build(patchwork[[2]])$layout$panel_scales_y[[1]]$range$range)

patchwork + plot_annotation(title = 'the two parameters') +
  plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = plot_text_size+2),
                                            plot.title=element_text(size = plot_text_size+2,hjust = 0.5)) &
  xlim(min(p_ranges_x), max(p_ranges_x)) & 
  ylim(min(p_ranges_y), max(p_ranges_y))

```


### Plot (different distributions)
The combined plot of the different distributions superimposed on each other.
```{r plot_posterior_2, fig.align="center", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# create the dataframe to pot using ggplot
df_plot <- data.frame(
  theta = theta,
  thetaprior = thetaprior
  )
df_plot_hist <- data.frame(
  postpredk = postpredk,
  priorpredk = priorpredk
)
# wide to long format
df_plot <- gather(df_plot, parameter, value, theta:thetaprior, factor_key=TRUE)
df_plot_hist <- gather(df_plot_hist, parameter, value, postpredk:priorpredk, factor_key=TRUE)

# plot
plot_top <- ggplot(df_plot, aes(x=value,fill=parameter)) + 
  geom_density()+
  scale_fill_manual(values = c(plot_color_6, plot_color_7))+
  xlab('theta')+ylab('density')+
  ggtitle(paste0('k=',stan_data$k,', n=',stan_data$n))+
  theme_classic()+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y =element_blank())+
  labs(fill = "")+guides(color = FALSE)+
  theme(plot.title = element_text(hjust=0.5))

plot_bottom <- ggplot(df_plot_hist, aes(x=value,fill=parameter,color=parameter)) + 
  geom_histogram(position="identity",binwidth = 1,alpha=0.5)+
  scale_color_manual(values = c(plot_color_3, plot_color_4))+
  scale_fill_manual(values = c(plot_color_6, plot_color_7))+
  xlab('success count')+ylab('mass')+
  #ggtitle(paste0('k=',stan_data$k,', n=',stan_data$n))+
  theme_classic()+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y =element_blank())+
  labs(fill = "")+guides(color = FALSE)+
  theme(plot.title = element_text(hjust=0.5))


patchwork <- plot_top / plot_bottom
patchwork + plot_annotation(title = 'different distributions') & 
  theme(plot.tag = element_text(size = plot_text_size+2),
        plot.title=element_text(size = plot_text_size+2,hjust = 0.5))
```


