---
title: "Exercise 5.3.3"
description: | 
  Individual exercise document
author: |
  | [Saurabh Steixner-Kumar](https://steixnerkumar.github.io/)
  | \footnotesize (social): [\large\faLinkedinSquare](https://www.linkedin.com/in/SteixnerKumar/) -  [\large\faGithub](https://github.com/SteixnerKumar) -  [\large\faTwitter](https://twitter.com/SteixnerKumar)
output: 
  html_document: 
    fig_caption: yes
    fig_height: 10
    fig_width: 17
    number_sections: yes
    toc: yes
    toc_float: true
  html_notebook: 
    fig_caption: yes
    fig_height: 10
    fig_width: 17
    number_sections: yes
    toc: yes
  pdf_document:
    latex_engine: xelatex 
    fig_caption: yes
    fig_height: 10
    fig_width: 17
    keep_tex: no
    number_sections: yes
    toc: yes
editor_options: 
  chunk_output_type: inline
# runtime: shiny
header-includes:
  - \usepackage{titling}
  - '\defaultfontfeatures{Extension = .otf}'
  - '\usepackage{fontawesome}'
  - \preauthor{\begin{flushright}}
  - \postauthor{\end{flushright}}
---

# Question
Exercise 5.3.3 Rare Disease Suppose you are testing a cheap instrument for detecting a rare medical condition. After 170 patients have been screened, the test results show that 157 did not have the condition, but 13 did. The expensive ground truth assessment subsequently revealed that, in fact, none of the patients had the condition. These data correspond to a = 0, b = 0, c = 13, d = 157. Apply the kappa graphical model to these data, and reach a conclusion about the usefulness of the cheap instrument. What is special about this data set, and what does it demonstrate about the Bayesian approach?

# Comments/Solution
The posterior mean for kappa is approximately .05, with a 95% credible interval (ETI) of approximately (0, .23). The data are noteworthy because the disease has never been observed, so there are two zero cells, and a zero column sum. This poses a challenge for frequentist estimators. In order to deal with the problem of zero counts a frequentist may add a “1” to each cell in the design, but this amounts to fabricating data. An attractive property of the Bayesian approach is that it is always possible to do the analysis.

The model used to calculate the required values and the plots is scripted below. Copy/pasting the given code will generate the same result on your own machine.


# Code
## libraries
The libraries required for the script and the plots.
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
# clears workspace
rm(list=ls()) 
#load libraries
library(rstan)
library(bayestestR)
library(tidyr)
library(ggplot2)
library(patchwork)
```


## Data
The data required for this particular stan model.
```{r libraries, data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
# data  initialization
# Rare Disease
y <- c(0, 0, 13, 157)
# to be passed on to Stan
stan_data <- list(y=y)
```


## Stan code
Stan code, that can be written in R as such or in a separate new file with stan extension.
```{r stan_code, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
write("// Stan code here in this section

// Kappa Coefficient of Agreement
data { 
  int<lower=0> y[4];
}
parameters {
  // Underlying Rates
  // Rate Objective Method Decides 'one'
  real<lower=0,upper=1> alpha;
  // Rate Surrogate Method Decides 'one' When Objective Method Decides 'one'
  real<lower=0,upper=1> beta;
  // Rate Surrogate Method Decides 'zero' When Objective Method Decides 'zero'
  real<lower=0,upper=1> gamma;
} 
transformed parameters {
  simplex[4] pi;
  real xi;
  real psi;
  real kappa;
  // Probabilities For Each Count
  pi[1] = alpha * beta;
  pi[2] = alpha * (1 - beta);
  pi[3] = (1 - alpha) * (1 - gamma);
  pi[4] = (1 - alpha) * gamma;
    
  // Derived Measures   
  // Rate Surrogate Method Agrees With the Objective Method
  xi = alpha * beta + (1 - alpha) * gamma ;
  // Rate of Chance Agreement
  psi = (pi[1] + pi[2]) * (pi[1] + pi[3]) + (pi[2] + pi[4]) * (pi[3] + pi[4]);  
  // Chance-Corrected Agreement
  kappa = (xi - psi) / (1 - psi);
}
model {
  alpha ~ beta(1, 1);  // could be removed
  beta ~ beta(1, 1);  // could be removed
  gamma ~ beta(1, 1);  // could be removed
  // Count Data     
  y ~ multinomial(pi);
} // ",

"5_3_3.stan")
```

## code in R to run stan
Running stan through R (with the required input parameters).
```{r running_stan, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hide'}

myinits <- list(
  list(alpha=.5, beta=.5, gamma=.5),  # chain 1 starting value
  list(alpha=.5, beta=.5, gamma=.5))  # chain 2 starting value

# parameters to be monitored:  
parameters <- c("kappa", "xi", "psi", "alpha", "beta", "gamma", "pi")

# The following command calls Stan with specific options.
# For a detailed description type "?stan".
mod_fit <- stan(file="5_3_3.stan",   
                data=stan_data, 
                init=myinits,  # If not specified, gives random inits
                pars=parameters,
                iter=2000, 
                chains=2, 
                thin=1,
                warmup=100,  # Stands for burn-in; Default = iter/2
                seed=123  # Setting seed; Default is random seed
)
```

# Outputs
## Model summary
In order of definition.
```{r model_summary, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# The commands below are useful for a quick overview:
print(mod_fit)  # a rough summary
#plot(mod_fit)   # a visual representation
```

```{r model_parameters, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
# get the data from the model
kappa   <- rstan::extract(mod_fit)$kappa
xi <- rstan::extract(mod_fit)$xi
psi   <- rstan::extract(mod_fit)$psi
a <- rstan::extract(mod_fit)$pi[,1]
b <- rstan::extract(mod_fit)$pi[,2]
c <- rstan::extract(mod_fit)$pi[,3]
d <- rstan::extract(mod_fit)$pi[,4]
#
# get the confidence interval
ci_hdi <- ci(kappa, ci=0.95,method = "HDI")
ci_eti <- ci(kappa, ci=0.95,method = "ETI")
```

## Plots

```{r plot_defaults, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
# Diagnostic plot
# stan_par(mod_fit, 'theta', chain = 0)

# First, some options to make the plot look better:
#
plot_color_1 <- adjustcolor("black",alpha.f=1)
plot_color_2 <- adjustcolor("white",alpha.f=1)
plot_color_3 <- adjustcolor("darkred",alpha.f=1)
plot_color_4 <- adjustcolor("darkgreen",alpha.f=1)
plot_color_5 <- adjustcolor("blue",alpha.f=0.8)
plot_color_6 <- adjustcolor("darkred",alpha.f=0.6)
plot_color_7 <- adjustcolor("darkgreen",alpha.f=0.6)
plot_text_size <- 32
plot_text_color <- adjustcolor("black",alpha.f=1)

```


### Plot (chains)
The initial movement of the chains are shown here (including the warmup phase). The two chains begin from the initial starting points of as defined in the input parameters of the stan model.
```{r plot_chain, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE, paged.print=TRUE}
# chain plot
x_limit <- c(0,200)
#y_limit <- c(0.6,1)
#
trace_1 <- traceplot(mod_fit, 'xi', include = TRUE, unconstrain = FALSE,inc_warmup = TRUE,show.legend = FALSE)+
  xlim(x_limit[1], x_limit[2])+
  #ylim(y_limit[1], y_limit[2])+
  scale_color_manual(values = c(plot_color_3, plot_color_4)) +
  theme(legend.position = "bottom")+
  theme_classic()+
  ylab('xi')+xlab('iterations')+
  #ggtitle(paste0('k=',stan_data$k,', n=',stan_data$n))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y = element_text(color = plot_text_color, size = plot_text_size))+
  theme(plot.title = element_text(hjust=0.5))
trace_2 <- traceplot(mod_fit, 'psi', include = TRUE, unconstrain = FALSE,inc_warmup = TRUE,show.legend = FALSE)+
  xlim(x_limit[1], x_limit[2])+
  #ylim(y_limit[1], y_limit[2])+
  scale_color_manual(values = c(plot_color_3, plot_color_4)) +
  theme(legend.position = "bottom")+
  theme_classic()+
  ylab('psi')+xlab('iterations')+
  #ggtitle(paste0('k=',stan_data$k,', n=',stan_data$n))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y = element_text(color = plot_text_color, size = plot_text_size))+
  theme(plot.title = element_text(hjust=0.5))
trace_3 <- traceplot(mod_fit, 'kappa', include = TRUE, unconstrain = FALSE,inc_warmup = TRUE)+
  xlim(x_limit[1], x_limit[2])+
  scale_color_manual(values = c(plot_color_3, plot_color_4)) +
  theme(legend.position = "bottom")+
  theme_classic()+
  ylab('kappa')+xlab('iterations')+
  #ggtitle(paste0('k=',stan_data$k,', n=',stan_data$n))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y = element_text(color = plot_text_color, size = plot_text_size))+
  theme(plot.title = element_text(hjust=0.5))
#
patchwork <- trace_1+trace_2
p_ranges_x <- c(ggplot_build(patchwork[[1]])$layout$panel_scales_x[[1]]$range$range,
                ggplot_build(patchwork[[2]])$layout$panel_scales_x[[1]]$range$range)

p_ranges_y <- c(ggplot_build(patchwork[[1]])$layout$panel_scales_y[[1]]$range$range,
                ggplot_build(patchwork[[2]])$layout$panel_scales_y[[1]]$range$range)
patchwork <- patchwork &
  xlim(min(p_ranges_x), max(p_ranges_x)) & 
  ylim(min(p_ranges_y), max(p_ranges_y))
#
patchwork <- patchwork/trace_3 
patchwork + plot_annotation(title = 'the three parameters') +
  plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = plot_text_size+2),
                                            plot.title=element_text(size = plot_text_size+2,hjust = 0.5))

```

### Plot (posterior)
The plot of the rate of agreement ($\xi$), rate of agreement by chance ($\psi$) and the chance-corrected agreement rate ($\kappa$) values per chain superimposed on each other. 
```{r plot_posterior, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE, paged.print=TRUE}
# Now let's plot a histogram/density for theta.
# density plot
x_limit <- c(0.6,1)
#y_limit <- c(0,20)
# Plot
dens_1 <- stan_dens(mod_fit, 'xi', alpha = 0.3, include = TRUE, unconstrain = FALSE,
            inc_warmup = FALSE, separate_chains = TRUE,show.legend = FALSE) +
  xlim(x_limit[1], x_limit[2])+
  #ylim(y_limit[1], y_limit[2])+
  scale_fill_manual(values = c(plot_color_3, plot_color_4))+
  xlab('xi')+ylab('density')+
  #ggtitle(paste0('k=',stan_data$k,', n=',stan_data$n))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y =element_blank())+
  theme(plot.title = element_text(hjust=0.5))
dens_2 <- stan_dens(mod_fit, 'psi', alpha = 0.3, include = TRUE, unconstrain = FALSE,
            inc_warmup = FALSE, separate_chains = TRUE,show.legend = FALSE) +
  xlim(x_limit[1], x_limit[2])+
  #ylim(y_limit[1], y_limit[2])+
  scale_fill_manual(values = c(plot_color_3, plot_color_4))+
  xlab('psi')+ylab('density')+
  #ggtitle(paste0('k=',stan_data$k,', n=',stan_data$n))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y =element_blank())+
  theme(plot.title = element_text(hjust=0.5))
dens_3 <- stan_dens(mod_fit, 'kappa', alpha = 0.3, include = TRUE, unconstrain = FALSE,
            inc_warmup = FALSE, separate_chains = TRUE) +
  scale_fill_manual(values = c(plot_color_3, plot_color_4))+
  xlab('kappa')+ylab('')+
  #ggtitle(paste0('k=',stan_data$k,', n=',stan_data$n))+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y =element_blank())+
  theme(plot.title = element_text(hjust=0.5))
#
patchwork <- dens_1+dens_2
p_ranges_x <- c(ggplot_build(patchwork[[1]])$layout$panel_scales_x[[1]]$range$range,
                ggplot_build(patchwork[[2]])$layout$panel_scales_x[[1]]$range$range)

p_ranges_y <- c(ggplot_build(patchwork[[1]])$layout$panel_scales_y[[1]]$range$range,
                ggplot_build(patchwork[[2]])$layout$panel_scales_y[[1]]$range$range)
patchwork <- patchwork &
  xlim(min(p_ranges_x), max(p_ranges_x)) & 
  ylim(min(p_ranges_y), max(p_ranges_y))
#
patchwork <- patchwork/dens_3 
patchwork + plot_annotation(title = 'the three parameters') +
  plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = plot_text_size+2),
                                            plot.title=element_text(size = plot_text_size+2,hjust = 0.5))


```

### Plot (posterior with CI)
The plot of the combined $\kappa$ value with confidence intervals. (HDI CI : blue; ETI CI : green; mean : dashed black)
```{r plot_ci, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE, paged.print=TRUE}
# the data
df_plot <- data.frame(
  kappa = kappa
  )
df_plot <- gather(df_plot, parameter, value, kappa, factor_key=TRUE)
# plot
plot_1 <- ggplot(df_plot, aes(x=value,fill=parameter)) +
  geom_density(show.legend = FALSE)+
  scale_fill_manual(values = c(plot_color_6))+
  xlab('kappa')+ylab('density')+
  #ggtitle(paste('dataset; a,b,c,d: ',y))+
  # HDI CI
  geom_vline(aes(xintercept=ci_hdi$CI_low), color=plot_color_5, size=3,show.legend=F) +
  geom_vline(xintercept=ci_hdi$CI_high, color=plot_color_5, size=3) +
  #geom_text(aes(x=ci_hdi$CI_high, label="HDI", y=2), color=plot_color_5, angle=-90, vjust = -0.9)+
  # ETI CI
  geom_vline(aes(xintercept=ci_eti$CI_low), color=plot_color_7, size=2,show.legend=F) +
  geom_vline(xintercept=ci_eti$CI_high, color=plot_color_7, size=2) +
  #geom_text(aes(x=ci_eti$CI_high, label="ETI", y=2), color=plot_color_7, angle=-90, vjust = 1.2)+
  geom_vline(xintercept=mean(kappa), color=plot_color_1, size=4, linetype='dashed') +
  theme_classic()+
  theme(legend.text = element_text(color = plot_text_color, size = plot_text_size),
        legend.title = element_text(color = plot_text_color, size = plot_text_size),
        legend.key.size = unit(3,"line"),
        axis.text.x = element_text(size = plot_text_size, angle = 0),
        axis.text.y = element_text(size = plot_text_size, angle = 0),
        axis.title.x = element_text(color = plot_text_color, size = plot_text_size),
        axis.title.y =element_blank())+
  labs(fill = "")+guides(color = FALSE)+
  theme(plot.title = element_text(hjust=0.5))
#
plot_1
```  

## Confidence interval value and the mean value
The confidence interval values.
```{r ci, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE, paged.print=TRUE}
# Confidence intervals
ci_hdi
ci_eti
```
The mean value
```{r mean, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE, paged.print=TRUE}
# mean
mean(kappa)
```
